Supplementary Material for 'The complex network of trophic interactions in

### 2023-07-12 MS_ToSubmit/MS_Burdwood_foodweb.tex
Schejter suggs.

  \begin{abstract}
  Globally, the total area of the world ocean designated under marine
  protection has increased in the recent decades. The majority of these
  Marine Protected Areas (MPAs) focus on the presence of particularly
  vulnerable, keystone, or charismatic species, large numbers of endemic
  species, and/or high biodiversity across taxonomic levels. In the
  sub-Antarctic region, the level of ocean protection is mainly
  associated to oceanic islands, except for the MPAs Namuncurá -
  Burdwood Bank I and II (MPA N-BB, 53º--55ºS and 56º--62ºW,
  respectively), which are associated to a submarine plateau and its
  southern adjacent deep slope region, respectively. Here, we present
  the first detailed analysis of the network of predator-prey
  interactions (food web) for the MPA N-BB ecosystem. We applied a
  network approach to characterise the food web in terms of complexity
  and structure, and identifying the species' role in such a framework.
  In terms of complexity, the MPA N-BB food web consisted of 1788
  predator-prey interactions and 380 species, with a link density of
  4.70 and a connectance of 0.01. In terms of structure, almost half of
  the consumers were omnivores (0.48), and the network displayed a
  small-world pattern. These findings suggest that the ecosystem might
  be vulnerable to external perturbations targeting highly connected
  species, although structural properties might provide resilience and
  resistance, resulting in a rearranged structure that preserves its
  original functions. Furthermore, we identified several species as
  important in terms of different aspects of trophic structure and
  functioning, and response to perturbations. We suggest that generalist
  species, mainly fishes, play a crucial role in the ecosystem's
  bentho-pelagic coupling process and that other species, besides the
  longtail southern cod \emph{Patagonotothen ramsayi} and the Fueguian
  sprat \emph{Sprattus fuegensis}, should be considered as relevant
  energy transfers for the ecosystem. Finally, we argue that the
  diversity of species, including both the benthic and pelagic habitats,
  is responsible for seccuring the connectivity within the food web
  against perturbations, therefore contributing to the structure and
  stability of the ecosystem.
  \end{abstract}
    \begin{keyword}
    Food web \sep Complexity \sep Structure \sep Burdwood Bank \sep 
    Southwest Atlantic
  \end{keyword}

@@

In the sub-Antarctic region, the level of ocean protection is mainly
associated to oceanic islands, such as the South Georgia and South
Sandwich, Bouvet, Prince Edward, and Macquarie islands (IUCN \&
UNEP-WCMC, 2023). Interestingly, the case of the MPAs Namuncurá -
Burdwood Bank I and II (MPA N-BB, 53º--55ºS 56º--62ºW), which is the
focus of this work, is unique since these MPAs are associated to a
submarine plateau and its southern adjacent deep slope region,
respectively (Falabella, 2017). In addition, such MPAs are part of a
network of protected areas in the sub-Antarctic area (jointly with MPA
Yaganes) that aims to protect and strengthen the functioning of this
southern region to maintain global ocean health.

Many of these MPAs focus on the presence of particularly vulnerable,
keystone, or charismatic species, large numbers (or proportions) of
endemic species, and/or high biodiversity across taxonomic levels (Hogg
et al., 2016). Indeed, the MPA N-BB was created to protect a sensitive
and biodiverse benthic habitat that was barely known (Falabella, 2017;
Schejter et al., 2016). The benthic community is featured by high
biomass of vulnerable and fragile species (mainly Porifera and Cnidaria)
that meet the characteristics of vulnerable marine ecosystems (Schejter
\& Albano, 2021), here defined as sites that present taxa with a minimum
biomass of 10 kg per 1200 m\textsuperscript{2} (Commission for the
Conservation of Antarctic Marine Living Resources (CCAMLR), 2009). Also,
it provides habitat to several small-sized species (López-Gappa et al.,
2018; Martin Sirito, 2019; Schejter \& Bremec, 2019), and has an
important role in the life history of fishes as a food source, refuge
and nursery area (Covatti Ale et al., 2022; Delpiani et al., 2020;
Fischer et al., 2022; Florencia et al., 2023; García Alonso et al.,
2018; Troccoli et al., 2020; Vazquez et al., 2018). The maintenance of
this singular community is related to local and regional oceanographic
processes, including the circulation of the rich Malvinas (Falkland)
current in the area (Guerrero et al., 1999; Piola \& Gordon, 1989) and
the upwelling and mixing phenomena (Matano et al., 2019). The input of
nutrients from the Malvinas (Falkland) current also supports a diverse
plankton community (Guinder et al., 2020). Overall, 811 benthic and
plankton species have been identified for the MPA N-BB ecosystem, where
349 were reported for the first time in the area in recent years
(Administración de Parques Nacionales, 2022). However, identifying the
main species involved in the maintenance of ecosystem services and
health as well as for management and conservation is essential.

Recently, the structure of the southwestern South Atlantic Ocean has
been proposed to be under a `wasp-waist' control, meaning that the
structure and dynamics of the ecosystem are regulated primarily by
mid-trophic level species (e.g., fishes, crustaceans) (Padovani et al.,
2012; Riccialdelli et al., 2020; Saporiti et al., 2015). In particular,
the ecosystem of the MPA N-BB shows a more pronounced `wasp-waist'
structure, meaning a shorter food-chain length and a greater trophic
overlap and redundancy, than other sub-Antarctic areas, such as the
continental shelf off Tierra del Fuego. The Fuegian sprat \emph{Sprattus
fuegensis} and longtail southern cod \emph{Patagonotothen ramsayi} are
considered the most plausible `wasp-waist' species (Riccialdelli et al.,
2020).

High-latitude marine ecosystems, such as the MPA N-BB, are considered
complex systems in terms of biodiversity and ecological interactions
(Cordone et al., 2020; Day et al., 2013; Kortsch et al., 2019; Trathan
et al., 2021). Although there is a robust knowledge about the complexity
considering the richness of the benthic and plankton communities in the
MPA N-BB ecosystem (Administración de Parques Nacionales, 2022; Guinder
et al., 2020; Schejter et al., 2016, 2020), a better understanding of
species interactions' complexity and structure is needed. This aspect
can be tackled by analysing one of the most frequent relationships
between species: the predator-prey interaction (Bascompte, 2009). The
sum of predator-prey or trophic interactions of a particular region is
referred to as a food web, representing the roadmap for matter and
energy flow in an ecosystem. In recent years, network approaches have
been successfully applied to study complex high-latitude marine
ecosystems, improving our knowledge on structure, functioning, and
response to environmental/anthropogenic changes (Cordone et al., 2018;
Funes et al., 2022; Kortsch et al., 2015; Marina et al., 2023). Among
anthropogenic threats, it is worth mentioning that contaminants like
mercury and microplastics have been recently reported as important
threats to the MPA N-BB region (Cossi et al., 2021; Di Mauro et al.,
2022; Fioramonti et al., 2022); also fishing vessels are allowed to
operate in the western section of the MPA N-BB (i.e.~Marine National
Reserve category), altering the stocks of commercially important fish
species (Administración de Parques Nacionales, 2022; Martínez et al.,
2021). Moreover, there is a potential hazard related to the effects of
offshore activities (exploration and explotation) to the west of the MPA
N-BB (Administración de Parques Nacionales, 2022).

In the present work, we present the first detailed analysis of the
network of predator-prey interactions, hereafter food web, for the MPA
N-BB ecosystem. For this, we applied a network approach to a highly
resolved food web. The objective was twofold: characterise the food web
in terms of complexity and structure, and identify the species' role in
such a network framework.

@@

The MPAs Namuncurá - Burdwood Bank I and II, created by National Laws
26.875 in 2013 and 27.490 in 2017, comprise a shallow submarine plateau
called Burdwood Bank (BB) and a deep slope that reaches 4000 m in depth,
N-BB I and N-BB II, respectively (Administración de Parques Nacionales,
2022; Tombesi et al., 2020) (Figure 1). They are located 150 km east of
Isla de los Estados and 200 km south of Malvinas/Falkland Islands. The
MPA N-BB I comprises nearly 34,000 km\textsuperscript{2} circumscribed
by the 200 m isobath, between 54º--55ºS and 56º--62ºW, with a slight
slope extended nearly 370 km east--west. Physical features in the BB are
fairly stable, with salinity averaging 34 all year round and temperature
ranging between 4 and 8ºC (Acha et al., 2004; Guerrero et al., 1999;
Piola \& Falabella, 2009). The BB is surrounded by steep flanks of up to
4000 m depth through which strong currents circulate (Matano et al.,
2019; Piola \& Gordon, 1989; Reta, 2014). The N-BB II includes such a
deep slope, protecting about 32,000 km\textsuperscript{2}. Intense
upwelling and mixing occur over it, entraining deep nutrient-rich waters
into the photic layer (Matano et al., 2019; Piola \& Falabella, 2009)
and resulting in a fairly homogeneous water column both spatially and
temporally (Glorioso \& Flather, 1995; Guerrero et al., 1999; Matano et
al., 2019).

@@

In order to build the network of predator-prey interactions, we reviewed
more than 150 references considering published articles, Ph.D.~theses,
public databases, and reports belonging to 16 research cruises conducted
in the MPAs N-BB I and II during 2014-2019. It is noteworthy that the
sampling effort was greater in the MPA N-BB I. Furthermore, we took into
account personal communications from experts belonging to the working
group of the study area
(https://www.pampazul.gob.ar/tag/banco-burdwood/). The diversity of the
authors' expertise contributing to the present study was a key factor in
enhancing the quality of the network, and inherently improved the
network representation. A list of the references used to build the
network is presented in Supplementary Material (Table S1).

Due to a lack of trophic data resolution for some species inhabiting the
study area, we followed the concept of trophic species, here defined as
aggregated groups of taxa. In most cases, we followed it when specific
data on species, in the taxonomic sense, were not available. In some
cases, we collapsed species when taxa shared the same set of predators
and prey (trophic similarity, Martinez (1991)), one of the aggregation
methods that better preserve food web functional properties (Gauzens et
al., 2013). In addition, for endemic species (e.g.~bryozoan
\emph{Burdwoodipora paguricola}) and other species with no trophic
studies so far, we inferred their feeding interactions applying a
conservative approach that assumes that the set of prey and predators
are at some point preserved in time. In those cases we gathered
information from upper taxonomic levels (i.e.~Genus, Family, Order,
Class, Phylum) as a good proxy variable (Morales-Castilla et al., 2015;
Pomeranz et al., 2019). Details about this can be found in Supplementary
Material (Table S2).

@@

The species-level analysis aims to describe the species' role in the
food web. For this, we considered the following properties: (1)
betweenness Btw; (2) closeness Cl; (3) trophic similarity TS; (4)
topological role TR; and (5) trophic level TL (Table 1). Topological
roles refer to the fact that food webs tend to naturally organize in
non-random, modular patterns, where modules are defined as a group of
species that interact more frequently among themselves than with species
that are not members of the module (Guimerà \& Nunes Amaral, 2005).
Species can play different roles in this respect, according to the
pattern of interactions within their own module and/or across modules.
We computed the topological role for each species, classified as module
hub, species with a relatively high number of interactions, but most
within its own module; module specialist, species with relatively few
interactions and most within its own module; module connector, species
with relatively few interactions mainly between modules; and network
connector, species with high connectivity between and within modules
(Guimerà \& Nunes Amaral, 2005).

@@

In terms of complexity, the MPA Namuncurá - Burdwood Bank food web (N-BB
I and II) consisted of 1788 predator-prey interactions and 380 species,
where 93\% of them were defined at the species taxonomical level (Figure
2, Table S2). The food web presented a link density (e.g., the average
number of interactions per species) of 4.70, and a connectance of 0.01.
Almost half of the consumers were omnivores (0.48), feeding on sources
at different trophic levels. The food web displayed a small-world
pattern, meaning that the path length was lower and the clustering
coefficient higher than the random networks (Table 2).

@@

\caption{\label{tab:table2}Network-level properties of the MPA Namuncurá - Burdwood Bank food web. CPL: Characteristic Path Length; CC: Clustering Coefficient; SW: Small-World pattern. See table 1 for definitions and ecological relevance.}
\centering
\begin{tabular}[t]{r|r|r|r|r|r|r|l}
\hline
\textbf{Species} & \textbf{Interactions} & \textbf{Density} & \textbf{Connectance} & \textbf{Omnivory} & \textbf{CPL} & \textbf{CC} & \textbf{SW}\\
\hline
380 & 1788 & 4.71 & 0.01 & 0.48 & 2.99 & 0.08 & True\\
\hline
\end{tabular}
\end{table}

The degree distribution of the food web showed an asymmetric frequency
in the number of interactions, where most of the species had a
relatively low number of interactions and few species concentrated most
of them (Figure 3A). The distribution of prey among predators showed
that most consumers fed on a low number of prey whereas few had multiple
prey (Figure 3B). The top-five predators in number of prey were:
yellowfin notothen \emph{Patagonotothen guntheri} (Notothenioid fish, 50
prey), rock cod \emph{Patagonotothen ramsayi} (Notothenioid fish, 49
prey), broad nose skate \emph{Bathyraja brachyurops} (Chondrichthyan, 33
prey), Patagonian toothfish \emph{Dissostichus eleginoides}
(Notothenioid fish, 30 prey), and graytail skate \emph{Bathyraja
griseocauda} (Chondrichthyan, 28 prey). Following the same distribution
pattern, few prey presented multiple predators (Figure 3C). The top-five
prey in number of predators were: Detritus (Non-living, 153 predators),
the three categories of Diatoms considered (benthic, centric and
pennate, 72.5 predators on average), and species of the genus
\emph{Euphausia} (Zooplankton, 46 predators). Finally, taking into
account the interactions within each functional group, most interactions
were concentrated in a few species (Figure 3D). The most evident species
were: \emph{Doryteuthis gahi} (Cephalopoda), \emph{Grimothea gregaria}
(Decapoda), \emph{Patagonotothen ramsayi}, \emph{Patagonotothen
guntheri} and \emph{Dissostichus eleginoides} (bentho-pelagic fish),
\emph{Sprattus fuegensis} and \emph{Micromesistius australis} (pelagic
fish), and species of \emph{Euphausia} and \emph{Themisto gaudichaudii}
(Zooplankton). Overall, there is an evident asymmetry in the
distribution of interactions among species at different levels in the
MPA N-BB food web.

@@

We found different relationships between the species trophic level (TL)
and the rest of the analysed species-level properties (Figure 4A-D). The
most evident significant relationship was with trophic similarity,
i.e.~the higher the species' TL, the lower the trophic similarity or the
higher the uniqueness in terms of trophic role (Figure 4C). Here it is
noteworthy to highlight those high-trophic level species (TL
\textgreater{} 3.1) with low values of trophic similarity:
\emph{Bathyraja macloviana} and \emph{Squalus acanthias}
(Chondrichthyans), \emph{Diplopteraster clarki} and \emph{Pteraster sp}
(echinoderms), \emph{Daption capense} and \emph{Eudyptes chrysocome}
(seabirds), Ziphiidae and \emph{Lagenorhynchus cruciger} (marine
mammals) (Table S3).

@@

Considering the topological role, `module specialist' species were the
most frequent and presented a wide TL range (1 - 4.78), as well as
`module hub' species (TL = 1 - 3.9); `module connector' was constrained
to mid-TLs (2 - 3.86); and `network connector', was represented by only
one trophic species: detritus (Figure 4D, see Figure S2 for species'
topological roles in a food web graph framework). Here it is important
to highlight the two latter topological roles because they are
responsible for linking modules and maintaining the connectivity of the
food web: 42 species (1 network connector + 41 module connectors) from
19 different functional groups with a TL range = 1 - 3.86. The 41
species with a module connector role represented these functional
groups: Amphipoda, Bivalvia, Brachiopoda, Bryozoa, Hydrozoa (as
`Cnidaria\_benthic'), Copepoda, Cumacea, Decapoda, Echinodermata, fish
(bentho-pelagic and demersal Osteichthyes, and Chondrychthyes),
Foraminifera, Polychaeta, Porifera, Pycnogonida (as `Benthos\_Misc') and
zooplankton (see Supplementary Material Table S3 for the identity of the
species).

@@

\caption{Species-level properties by trophic level: (A) betweenness, (B) closeness, (C) trophic similarity, and (D) topological role. Each point represents a species. Linear regressions for betweenness ($y = 74.82x - 116.82, R^2 = 0.05, p-value < 0.01$), closeness ($y = 1.04e-05x - 9.25e-4, R^2 = 0.004, p-value = 0.11$) and trophic similarity ($y = -0.02x + 0.1 , R^2 = 0.07, p-value < 0.01$). Note that for panels A, B and C only species with TLs equal or greater than 2 were considered.}\label{fig:figure4}
\end{figure}

@@

The food web of the MPA N-BB ecosystem analysed in this study is one of
the most highly-resolved networks of trophic interactions ever studied,
not only for a high-latitude marine ecosystem but also for any marine
protected area worldwide to our knowledge. It is of paramount importance
to consider the complexity of species interactions in order to gain
insights into the structure and functioning of the ecosystem, since the
aggregation of species might mask food web properties and produce type
II errors (false positive) (Gauzens et al., 2013; Martinez, 1993).

>>>>

  \begin{abstract}
  Globally, the total area of the world ocean designated under marine
  protection has increased in the recent decades. The majority of these
  Marine Protected Areas (MPAs) focus on the presence of particularly
  vulnerable, keystone, or charismatic species, large numbers of endemic
  species, and/or high biodiversity across taxonomic levels. In the
  sub-Antarctic region, the level of ocean protection is mainly
  associated to oceanic islands, except for the MPAs Namuncurá -
  Burdwood Bank I and II (MPA N-BB, \textasciitilde53º--55ºS and
  \textasciitilde56º--62ºW), which are associated to a submarine plateau
  and its southern adjacent deep slope region, respectively. Here, we
  present the first detailed analysis of the network of predator-prey
  interactions (food web) for the MPA N-BB ecosystem. We applied a
  network approach to characterise the food web in terms of complexity
  and structure, and identifying the species' role in such a framework.
  In terms of complexity, the MPA N-BB food web consisted of 1788
  predator-prey interactions and 379 species, with a link density of
  4.72 and a connectance of 0.01. In terms of structure, almost half of
  the consumers were omnivores (0.48), and the network displayed a
  small-world pattern. These findings suggest that the ecosystem might
  be vulnerable to external perturbations targeting highly connected
  species, although other properties might provide resilience and
  resistance, resulting in a rearranged structure that preserves its
  original functions. Furthermore, we identified several species as
  important in terms of different aspects of trophic structure and
  functioning, and response to perturbations. We suggest that generalist
  species, mainly fishes, play a crucial role in the ecosystem's
  bentho-pelagic coupling process and that other species, besides the
  longtail southern cod \emph{Patagonotothen ramsayi} and the Fueguian
  sprat \emph{Sprattus fuegensis}, should be considered as relevant
  energy transfers for the ecosystem. Finally, we argue that the
  diversity of species, including both the benthic and pelagic habitats,
  is responsible for seccuring the connectivity within the food web
  against perturbations, therefore contributing to the structure and
  stability of the ecosystem.
  \end{abstract}
    \begin{keyword}
    Food web \sep Complexity \sep Structure \sep Burdwood Bank \sep 
    Southwest Atlantic
  \end{keyword}

@@

In the sub-Antarctic region, the level of ocean protection is mainly
associated to oceanic islands, such as the South Georgias and South
Sandwich, Bouvet, Prince Edward, and Macquarie islands (IUCN \&
UNEP-WCMC, 2023). Interestingly, the case of the MPAs Namuncurá -
Burdwood Bank I and II (MPA N-BB, \textasciitilde53º--55ºS
\textasciitilde56º--62ºW), which is the focus of this work, is unique
since these MPAs are associated to a submarine plateau and its southern
adjacent deep slope region, respectively (Falabella, 2017; Schejter et
al., 2020). In addition, such MPAs are part of a network of protected
areas in the sub-Antarctic area (jointly with MPA Yaganes) that aims to
protect this southern region in order to contribute to global ocean
health.

Many of these MPAs focus on the presence of particularly vulnerable,
keystone, or charismatic species, large numbers (or proportions) of
endemic species, and/or high biodiversity across taxonomic levels (Hogg
et al., 2016). Indeed, the MPA N-BB was created to protect a potentially
sensitive and biodiverse benthic habitat that was only barely known
(Falabella, 2017; Schejter et al., 2016). The benthic community is
featured by high biomass of vulnerable and fragile species (mainly
Porifera, Bryozoa and Cnidaria) that considered with their environment
meet the characteristics of vulnerable marine ecosystems (Schejter \&
Albano, 2021), here defined as sites that present densities of Indicator
Taxa of \textgreater{} 10 kg per 1200 m\textsuperscript{2} (Commission
for the Conservation of Antarctic Marine Living Resources (CCAMLR),
2009). Also, the benthic realm provides habitat to several small-sized
species (López-Gappa et al., 2018; Martin Sirito, 2019; Schejter \&
Bremec, 2019), and has an important role in the life history of fishes
as a food source, refuge and nursery area (Covatti Ale et al., 2022;
Delpiani et al., 2020; Fischer et al., 2022; Florencia et al., 2023;
García Alonso et al., 2018; Matusevich, 2022; Troccoli et al., 2020;
Vazquez et al., 2018). The maintenance of this singular community is
related to local and regional oceanographic processes, including the
circulation of the rich Malvinas (Falkland) current in the area
(Guerrero et al., 1999; Piola \& Gordon, 1989) and the upwelling and
mixing phenomena (Matano et al., 2019). The input of nutrients from the
Malvinas (Falkland) current also supports a diverse plankton community
(Guinder et al., 2020).

Overall, 811 benthic and plankton species have been identified for the
MPA N-BB ecosystem, where 349 were reported for the first time in the
area in recent years (Administración de Parques Nacionales, 2022).
Identifying the main species involved in the maintenance of ecosystem
services and health as well as for management and conservation is
essential. Recently, the structure of the southwestern South Atlantic
Ocean has been proposed to be under a `wasp-waist' control, meaning that
the structure and dynamics of the ecosystem are regulated primarily by
mid-trophic level species (e.g., fishes, crustaceans) (Padovani et al.,
2012; Riccialdelli et al., 2020; Saporiti et al., 2015). In particular,
the ecosystem of the MPA N-BB shows a more pronounced `wasp-waist'
structure, meaning a shorter food chain and a greater trophic overlap
and redundancy, than other sub-Antarctic areas, such as the continental
shelf off Tierra del Fuego. The Fuegian sprat \emph{Sprattus fuegensis}
and longtail southern cod \emph{Patagonotothen ramsayi} are considered
the most plausible `wasp-waist' species (Riccialdelli et al., 2020).

High-latitude marine ecosystems, such as the MPA N-BB, are complex
systems in terms of biodiversity and ecological interactions (Cordone et
al., 2020; Day et al., 2013; Kortsch et al., 2019; Trathan et al.,
2021). Although there is a robust knowledge about the complexity
considering the richness of the benthic and plankton communities in the
MPA N-BB ecosystem (Administración de Parques Nacionales, 2022; Guinder
et al., 2020; Schejter et al., 2016, 2020), a better understanding of
species interactions' complexity and structure is needed. This aspect
can be tackled by analysing one of the most frequent relationships
between species: the predator-prey interaction (Bascompte, 2009). The
sum of predator-prey or trophic interactions of a particular region is
referred to as a food web, representing the roadmap for matter and
energy flow in an ecosystem. In recent years, network approaches have
been successfully applied to study complex high-latitude marine
ecosystems, improving our knowledge on structure, functioning, and
response to environmental/anthropogenic changes (Cordone et al., 2018;
Funes et al., 2022; Kortsch et al., 2015; Marina et al., 2023). Among
anthropogenic threats, it is worth mentioning that contaminants like
mercury and microplastics have been recently reported as important
threats to the MPA N-BB region (Cossi et al., 2021; Di Mauro et al.,
2022; Fioramonti et al., 2022); also fishing vessels are allowed to
operate in the western section of the MPA N-BB (i.e.~Marine National
Reserve category), altering the stocks of commercially important fish
species (Administración de Parques Nacionales, 2022; Martínez et al.,
2021). Moreover, there is a potential hazard related to the effects of
offshore activities (exploration and explotation) to the west of the MPA
N-BB (Administración de Parques Nacionales, 2022).

In the present work, we present the first detailed analysis of the
network of predator-prey interactions, hereafter food web, for the MPA
N-BB ecosystem. For this, we applied a network approach to a highly
resolved food web. The objective was twofold: characterise the food web
in terms of complexity and structure, and identify the species' role in
the network.

@@

The MPAs Namuncurá - Burdwood Bank I and II, created by National Laws
26.875 in 2013 and 27.490 in 2017, comprise a shallow submarine plateau
called Burdwood Bank (BB) and a deep slope that reaches 4000 m in depth,
N-BB I and N-BB II, respectively (Administración de Parques Nacionales,
2022; Tombesi et al., 2020) (Figure 1). They are located 150 km east of
Isla de los Estados and 200 km south of Malvinas/Falkland Islands. The
MPA N-BB I comprises nearly 28,900 km\textsuperscript{2} circumscribed
by the 200 m isobath, between \textasciitilde54º--55ºS and
\textasciitilde56º--62ºW, with a slight slope extended nearly 370 km
east--west. Physical features in the BB are fairly stable, with salinity
averaging 34 all year round and temperature ranging between 4 and 8ºC
(Acha et al., 2004; Guerrero et al., 1999; Piola \& Falabella, 2009).
The BB is surrounded by steep flanks of up to 4000 m depth through which
strong currents circulate (Matano et al., 2019; Piola \& Gordon, 1989;
Reta, 2014). The N-BB II includes such a deep slope, protecting about
32,000 km\textsuperscript{2} (\textasciitilde55º-56ºS,
\textasciitilde58º-62ºW). Intense upwelling and mixing occur in relation
with the slope, entraining deep nutrient-rich waters into the photic
layer (Matano et al., 2019; Piola \& Falabella, 2009) and resulting in a
fairly homogeneous water column both spatially and temporally (Glorioso
\& Flather, 1995; Guerrero et al., 1999; Matano et al., 2019).

@@

In order to build the network of predator-prey interactions, we reviewed
more than 170 references considering published articles, Ph.D.~theses,
public databases, and reports belonging to 16 research cruises conducted
in the MPAs N-BB I and II during 2014-2019. It is noteworthy that the
sampling effort was greater in the MPA N-BB I. Furthermore, we took into
account personal communications from experts belonging to the working
group of the study area
(https://www.pampazul.gob.ar/tag/banco-burdwood/). The diversity of the
authors' expertise contributing to the present study was a key factor in
enhancing the quality of the network, and inherently improved the
network representation. A list of the references used to build the
network is presented in Supplementary Material (Table S1).

Due to a lack of trophic data resolution for some species inhabiting the
study area, we followed the concept of trophic species, here defined as
follows: taxa collapsed into a single node in the network. In most
cases, we followed this concept when specific data on species, in the
taxonomic sense, were not available. In some cases, we collapsed species
when taxa shared the same set of predators and prey (trophic similarity,
Martinez (1991)), one of the aggregation methods that better preserve
food web functional properties (Gauzens et al., 2013). In addition, for
endemic species (e.g.~bryozoan \emph{Burdwoodipora paguricola}) and
other species with no trophic studies so far, we inferred their feeding
interactions applying a conservative approach that assumes that the set
of prey and predators are at some point preserved in time. In those
cases we gathered information from upper taxonomic levels (i.e.~Genus,
Family, Order, Class, Phylum) as a good proxy variable (Morales-Castilla
et al., 2015; Pomeranz et al., 2019). Details about this can be found in
Supplementary Material (Table S2). Furthermore, we considered non-living
food sources, such as detritus and necromass, as prey species in the
food web context.

@@

The species-level analysis aims to describe the species' role in the
food web. For this, we considered the following properties: betweenness
Btw, closeness Cl, trophic similarity TS, topological role TR, and
trophic level TL (Table 1). Topological roles refer to the fact that
food webs tend to naturally organize in non-random, modular patterns,
where modules are defined as a group of species that interact more
frequently among themselves than with species that are not members of
the module (Guimerà \& Nunes Amaral, 2005). Species can play different
roles in this respect, according to the pattern of interactions within
their own module and/or across modules. We computed the topological role
for each species, classified as module hub, species with a relatively
high number of interactions, but most within its own module; module
specialist, species with relatively few interactions and most within its
own module; module connector, species with relatively few interactions
mainly between modules; and network connector, species with high
connectivity between and within modules (Guimerà \& Nunes Amaral, 2005).

@@

In terms of complexity, the MPA Namuncurá - Burdwood Bank food web
consisted of 1788 predator-prey interactions and 379 species, where 93\%
of them were defined at the species taxonomical level (Figure 2, Table
S2). The food web presented a link density (e.g., the average number of
interactions per species) of 4.72, and a connectance of 0.01. Almost
half of the consumers were omnivores (0.48), feeding on sources at
different trophic levels. The food web displayed a small-world pattern,
meaning that the path length was lower and the clustering coefficient
higher than the random networks (Table 2).

@@

\caption{\label{tab:table2}Network-level properties of the MPA Namuncurá - Burdwood Bank food web. CPL: Characteristic Path Length; CC: Clustering Coefficient; SW: Small-World pattern. See table 1 for definitions and ecological relevance.}
\centering
\begin{tabular}[t]{r|r|r|r|r|r|r|l}
\hline
\textbf{Species} & \textbf{Interactions} & \textbf{Density} & \textbf{Connectance} & \textbf{Omnivory} & \textbf{CPL} & \textbf{CC} & \textbf{SW}\\
\hline
379 & 1788 & 4.72 & 0.01 & 0.49 & 2.99 & 0.08 & True\\
\hline
\end{tabular}
\end{table}

The degree distribution of the food web showed an asymmetric frequency
in the number of interactions, where most of the species had a
relatively low number of interactions and few species concentrated most
of them (Figure 3A). The distribution of prey among predators showed
that most consumers fed on a low number of prey whereas few had multiple
prey (Figure 3B). The top-five predators in number of prey were:
yellowfin notothen \emph{Patagonotothen guntheri} (Notothenioid fish, 50
prey), rock cod \emph{Patagonotothen ramsayi} (Notothenioid fish, 49
prey), broad nose skate \emph{Bathyraja brachyurops} (Chondrichthyan, 33
prey), Patagonian toothfish \emph{Dissostichus eleginoides}
(Notothenioid fish, 30 prey), and graytail skate \emph{Bathyraja
griseocauda} (Chondrichthyan, 28 prey). Following the same distribution
pattern, few prey presented multiple predators (Figure 3C). The top-five
prey (or food sources) in number of predators were: Detritus
(Non-living, 153 predators), the three categories of Diatoms considered
(benthic, centric and pennate, 72.5 predators on average), and species
of the genus \emph{Euphausia} (Zooplankton, 46 predators). Finally,
taking into account the interactions within each functional group, most
interactions were concentrated in a few species (Figure 3D). The most
evident species were: \emph{Doryteuthis gahi} (Cephalopoda),
\emph{Grimothea (=Munida) gregaria} (Decapoda), \emph{Patagonotothen
ramsayi}, \emph{Patagonotothen guntheri} and \emph{Dissostichus
eleginoides} (bentho-pelagic fish), \emph{Sprattus fuegensis} and
\emph{Micromesistius australis} (pelagic fish), and species of
\emph{Euphausia} and \emph{Themisto gaudichaudii} (Zooplankton).
Overall, there is an evident asymmetry in the distribution of
interactions among species at different levels in the MPA N-BB food web.

@@

We found different relationships between the species trophic level (TL)
and the rest of the analysed species-level properties (Figure 4A-D). The
most evident significant relationship was with trophic similarity,
i.e.~the higher the species' TL, the lower the trophic similarity or the
higher the uniqueness in terms of trophic role (Figure 4C). Here it is
noteworthy to highlight those high-trophic level species (TL
\textgreater{} 3.1) with low values of trophic similarity:
\emph{Bathyraja macloviana} and \emph{Squalus acanthias}
(Chondrichthyans), \emph{Diplopteraster clarki} and \emph{Pteraster} sp.
(echinoderms), \emph{Daption capense} and \emph{Eudyptes chrysocome}
(seabirds), Ziphiidae and \emph{Lagenorhynchus cruciger} (marine
mammals) (Table S3).

@@

Considering the topological role, `module specialist' species were the
most frequent and presented a wide TL range (1 - 4.78), as well as
`module hub' species (TL = 1 - 3.92); `module connector' was constrained
to mid-TLs (2 - 3.86); and `network connector', was represented by only
one trophic species: detritus (Figure 4D, see Figure S2 for species'
topological roles in a food web graph framework). Here it is important
to highlight the two latter topological roles because they are
responsible for linking modules and maintaining the connectivity of the
food web: 42 species (1 network connector + 41 module connectors) from
19 different functional groups with a TL range = 1 - 3.86. The 41
species with a module connector role represented these functional
groups: Amphipoda, Bivalvia, Brachiopoda, Bryozoa, Hydrozoa (as
`Cnidaria\_benthic'), Copepoda, Cumacea, Decapoda, Echinodermata, fish
(bentho-pelagic and demersal Osteichthyes, and Chondrychthyes),
Foraminifera, Polychaeta, Porifera, Pycnogonida (as `Benthos\_Misc') and
zooplankton (see Supplementary Material Table S3 for the identity of the
species).

@@

\caption{Species-level properties by trophic level: (A) betweenness, (B) closeness, (C) trophic similarity, and (D) topological role. Each point represents a species. Linear regressions for betweenness ($y = 74.97x - 117.35, R^2 = 0.05, p-value < 0.01$), closeness ($y = 9.33e-06x - 9.31e-4, R^2 = 0.003, p-value = 0.15$) and trophic similarity ($y = -0.02x + 0.11 , R^2 = 0.07, p-value < 0.01$). Note that for panels A, B and C only species with TLs equal or greater than 2 were considered.}\label{fig:figure4}
\end{figure}

@@

The food web of the MPA N-BB ecosystem analysed in this study is one of
the most highly-resolved networks of trophic interactions ever studied,
not only for a high-latitude open-ocean ecosystem but also for any
marine protected area worldwide to our knowledge. It is of paramount
importance to consider the complexity of species interactions in order
to gain insights into the structure and functioning of the ecosystem,
since the aggregation of species might mask food web properties and
produce type II errors (false positives) (Gauzens et al., 2013;
Martinez, 1993).

Food web connectance is a feature that resumes the complexity of the
network, but more importantly, it is an emergent property of pairwise
species interactions (Poisot \& Gravel, 2014). It contains information
regarding how interactions within an ecological network are distributed
and predicts reasonably well key dynamical properties of ecological
networks (Dunne et al., 2002c). Complex marine food webs (i.e.~with more
than 25 trophic species) show connectance values ranging from 0.01 -
0.27 (Marina, Saravia, et al., 2018). In particular, food webs from
high-latitude regions tend to exhibit a connectance closer to the
minimum (between 0.01 and 0.05) (Kortsch et al., 2015; Rodriguez et al.,
2022; Santana et al., 2013). Whether food webs display a low or a high
connectance helps to better comprehend ecosystem's synthetic properties
like robustness. In this sense, empirical analyses support the notion
that highly-connected ecological networks are robust against external
perturbations such as the introduction of new (e.g., invasive) species
(Smith-Ramesh et al., 2017) as well as species removal (e.g., local
extinction) (Dunne et al., 2002a; Montoya \& Solé, 2003). The
connectance of the food web of the MPA Namuncurá - Burdwood Bank (0.01)
is one of the lowest reported so far for these regions; in particular,
it appears to be much lower than that of Beagle Channel (0.05), an
adjacent coastal area (Rodriguez et al., 2022).

### 2023-07-12 MS_ToSubmit/SupMaterial.Rmd
Schejter suggs.

# Load data
load("../data/cleaned-data_jun23.rda")

@@

# Load data
load("../data/cleaned-data_jun23.rda")
load("../data/foodweb-data_jun23.rda")

@@

# Load data
load("../results/summary_results_jun23.rda")

>>>>

# Load data
load("../data/cleaned-data_jul23.rda")

@@

# Load data
load("../data/cleaned-data_jul23.rda")
load("../data/foodweb-data_jul23.rda")

@@

# Load data
load("../results/summary_results_jul23.rda")
